---
title: "HW1"
author: "David Zhao"
date: "9/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r}
#loading data and summary statistics 
data("penguins", package = "palmerpenguins")
summary(penguins)
#dataframe dimension
nrow(penguins)
ncol(penguins)
mean(na.omit(penguins$flipper_length_mm))
```

The penguins data set contains 344 rows and 8 columns. It includes 8 different characteristics about penguins from island in Palmer Archipelago, including species, which island it was located in, flipper length in mm, sex, body mass in grams and etc. There were a total of 11 rows with missing data. The mean flipper length, after removing the rows with missing values, was 200.9 mm, and the mean body mass was  4202 grams.

```{r}
# Figure 1: Flipper length vs. bill length by species
library(ggplot2)
ggplot(penguins,aes(x=bill_length_mm,y=flipper_length_mm,color=species))+geom_point()
ggsave("Figure1.png")
```

## Question 2

```{r,message=FALSE}
#generate 10 radom integers for factor
set.seed(100)
y = sample(1:3,10,replace=TRUE)
suppressMessages(library(tidyverse))
problem2_df=tibble(
  x = rnorm(10),
  whether_positive = x>0,
  character_var = character(length=10L),
  factor_var = factor(y,labels=c("level1","level2","level3")),
)
```


```{r,warning=FALSE}
# mean of each variable in df 
mean(pull(problem2_df,var = x))
mean(pull(problem2_df,var = whether_positive))
mean(pull(problem2_df,var = character_var))
mean(pull(problem2_df,var = factor_var))
```

The mean value for the numeric column was 0.265 and the mean for the logical column was 0.6. The mean function did not work for the character or the factor data type, and it returned NA instead.

```{r,eval=FALSE}
as.numeric(problem2_df$whether_positive)
as.numeric(problem2_df$character_var)
as.numeric(problem2_df$factor_var)
```

The as.numeric function converts the logical variables into values of 0s and 1s, which explained the value of 0.6 obtained previously. For the character variables, as.numeric returned a series of NAs. The as.numeric converts the three "levels" in the factor variables into three distinct numeric values.